The Kapowski Chronicles: A Complete, Volumetric Cortical Thickness Open Source Pipeline with Evaluation on Public Data

### 2013-05-29 KapowskiChronicles/methods.tex
ENH:  Lil John

\begin{figure*}
  \centering
  \includegraphics[width=170mm]{Figures/Kapowski_pipeline2.pdf}
  \caption{Illustration of the main components of the ANTs processing 
  workflow containing all elements for determining cortical thickness. 
  Not shown is the optional subject to template registration.}
  \label{fig:pipeline}
\end{figure*}

@@

  The warped template probability map is thresholded at 0.5 and the resulting mask is dilated
with a radius of 2.  Atropos is used to generate an initial 3-tissue segmentation estimate within the mask
region.  Each of the three tissue masks undergo specific morphological operations which are then
combined to create a brain extraction mask for use in the rest of the
cortical thickness workflow.  \textcolor{blue}{does there need to be a
  bit more technical detail here?  while you can refer to the script,
  why perform these operations?  there are numerous references, most
  germane probably some recent stuff from J Prince ( i think ) and the
  freesurfer watershed approach which came from ... i forget ... maybe
  one of the french groups.}

A comparison using open access brain data with publicly available
brain extraction algorithms including AFNI's \verb#3dIntracranial#
\citep{ward1999}, FSL's \verb#BET2# \citep{smith2002}, Freesurfer's
\verb#mri_watershed# \citep{segonne2004}, and BrainSuite
\citep{dogdas2005} demonstrated that our combined
registration/segmentation approach \citep{avants2010a} performs at the
top level alongside BrainSuite (tuned) and FreeSurfer.
\textcolor{blue}{ok you have the segonne ref here ... }

@@

@@

@@

\paragraph{Oasis}

>>>>

\begin{figure*}
  \centering
  \includegraphics[width=180mm]{Figures/Kapowski_pipeline2.pdf}
  \caption{Illustration of the main components of the ANTs processing 
  workflow containing all elements for determining cortical thickness. 
  We also included the domain of operations for select scripts also available
  in ANTs.
  Not shown is the optional subject to template registration.}
  \label{fig:pipeline}
\end{figure*}

@@

  The warped template probability map is thresholded at 0.5 and the resulting mask is dilated
with a radius of 2.  Atropos is used to generate an initial 3-tissue segmentation estimate within the mask
region.  Each of the three tissue masks undergo specific morphological operations which are then
combined to create a brain extraction mask for use in the rest of the
cortical thickness workflow.

A comparison using open access brain data with publicly available
brain extraction algorithms including AFNI's \verb#3dIntracranial#
\citep{ward1999}, FSL's \verb#BET2# \citep{smith2002}, Freesurfer's
\verb#mri_watershed# \citep{segonne2004}, and BrainSuite
\citep{dogdas2005} demonstrated that our combined
registration/segmentation approach \citep{avants2010a} performs at the
top level alongside BrainSuite (tuned) and FreeSurfer.

@@

In total, we processed 1202 T1-weighted images from four different
public data sets to obtain the corresponding cortical thickness maps.                           
The four data sets are described below:

@@

\paragraph{Oasis}
The Open Access Series of Imaging Studies (OASIS)%
\footnote{
http://www.oasis-brains.org/
}
 project consists of 416
(160 males and 256 females) T1-weighted images spanning 
the ages between 18 and 96 with 100 of these subjects having been 
diagnosed with very mild to moderate Alzheimer's disease.  
All subjects are right-handed.  Additional processing is included
with the distribution including bias corrected and segmented images
(gray, white and CSF matters).

\paragraph{NKI}
In support of open science, the 1,000 Functional Connectomes Project%
\footnote{ 
http://fcon\_1000.projects.nitrc.org
}
was initiated on December 11, 2009 by various members of the MRI community
seeking to form collaborative partnerships with imaging institutions for 
sharing well-documented multimodal image sets accompanied by phenotypic data.
This has resulted in the The Nathan Klein Institute (NKI)/Rockland sample
consisting of (at the time of download, September 22, 2012), 186 T1-weighted
images (87 females and 99 males).

\paragraph{Kirby}
The Multi-Modal MRI Reproducibility Resource%
\footnote{
http://www.nitrc.org/projects/multimodal/
}, 
or more informally the Kirby
data set, was originally described in \cite{landman2011} consisting of 
21 subjects (10 female and 11 male) and featuring a rich multiple modality and 
repeated acquisition schedule.

@@

### 2013-05-29 KapowskiChronicles/paper.tex
ENH:  Lil John

>>>>

\section{Comments from Brian}
\begin{itemize}
  \item contributions
  \begin{itemize}
    \item establish population variability internationally ... reference thickness values .... i.e. look up table for thickness results ....
    \item 7 sites 
    \item report novel gender diff over age 
    \item play up large - scale reliability  multisite 
    \item largest N on public multisite data?  reference peele ...
    \item quantify site effect .... is it low?
    \item allows us to quantify gender differences over a large age range ....
  \end{itemize}
  \item misc.
    \begin{itemize}
      \item figures  
      \item ask peele for feedback?
      \item questions - can we get by on large N and multisite alone 
      \item do we need to establish that we are more reliable than FSurf or FSL or SPM?
      \item we set the bar high by doing X number of images
      \item in discussion, say that future work will quantify SPM ....  not nec. by us. 
    \end{itemize}
\end{itemize}
